---
title: "supplemental_draft"
output: html_document
date: "2026-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggthemes)
library(ggridges)
library(scales)
library(tidyboot)
library(xtable)
library(papaja)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(weights)
```

```{r}
most_typicals <- read_csv(here("../data/all_corpora_most_typical_adjs.csv")) %>%
  filter(!is.na(adjective)) %>%
  mutate(adj_noun = paste(adjective, noun, sep = " "))

most_typicals %>%
  mutate(most_common_feature_rating = if_else(is.na(most_common_feature_rating), 0, most_common_feature_rating)) %>%
  ggplot(aes(x = most_common_feature_rating)) +
  geom_histogram()

most_typical_adj_nouns <- most_typicals %>%
  filter(most_common_feature_rating >= 4)
```

```{r read-data-judgments}
coded_data <- read_csv(here("../data/ldp_cabnc_data.csv"))

all_subjs <- coded_data %>%
  count(subid)

keep_subjs <- coded_data %>%
  filter(is.na(adjective) & is.na(noun) & rating == 5)

subj_data <- coded_data %>%
  filter(!(is.na(adjective) & is.na(noun))) %>%
  filter(subid %in% keep_subjs$subid) 

incomplete_subs <- coded_data %>%
  group_by(subid) %>%
  summarise(ntrials = n()) %>%
  filter(ntrials != 36)

nonsense_pairs <- subj_data %>%
  group_by(adjective, noun) %>%
  tally(rating == 8) %>%
  filter(n >= 2) %>%
  mutate(adj_noun_phrase = paste(adjective, noun, sep = " "))

subj_data <- subj_data %>%
  mutate(adj_noun_phrase = paste(adjective, noun, sep = " ")) %>%
  filter(!(adj_noun_phrase %in% nonsense_pairs$adj_noun_phrase),
         rating != 8)

mean_ratings <- subj_data %>%
  group_by(adjective, noun) %>%
  summarise(mean_typ = mean(rating)) %>%
  ungroup()

subj_reliability <- subj_data %>%
  group_by(adjective, noun) %>%
  reframe(rating = rating, rating_n = paste0("rating_", 1:n())) %>%
  pivot_wider(values_from = "rating", names_from = "rating_n") %>%
  rowwise() %>%
  mutate(mean_1 = mean(c(rating_2, rating_3, rating_4, rating_5, rating_6), na.rm = TRUE),
         mean_2 = mean(c(rating_1, rating_3, rating_4, rating_5, rating_6), na.rm = TRUE),
         mean_3 = mean(c(rating_1, rating_2, rating_4, rating_5, rating_6), na.rm = TRUE),
         mean_4 = mean(c(rating_1, rating_2, rating_3, rating_5, rating_6), na.rm = TRUE),
         mean_5 = mean(c(rating_1, rating_2, rating_3, rating_4, rating_6), na.rm = TRUE),
         mean_6 = mean(c(rating_1, rating_2, rating_3, rating_4, rating_5), na.rm = TRUE)) %>%
  ungroup()

```

```{r read-corpora}
all_corpora <- read_csv(here("../data/all_corpora_adj_noun_utts_redacted.csv")) %>%
  mutate(adj_noun = paste(adj_lower, noun_lemma, sep = " ")) %>%
  filter(adj_noun %in% most_typicals$adj_noun)

orig_tokens <- read_csv(here("../data/ldp_cabnc_pairs_all_info.csv"))

ldp <- all_corpora %>%
  filter(dataset != "cabnc_utts_udpipe")

all_corpus_counts <- all_corpora %>%
  left_join(orig_tokens, by = c("adj_token" = "adj", "noun_token" = "noun_orig")) %>%
  rename(adjective = adj_token) %>%
  count(adjective, noun) %>%
  ungroup()

cabnc <- all_corpora %>%
  filter(dataset == "cabnc_utts_udpipe") %>%
  left_join(orig_tokens, by = c("adj_token" = "adj", "noun_token" = "noun_orig")) %>%
  rename(adjective = adj_token) %>%
  select(doc_id, noun_token, adjective, noun, prenominal, 
         dataset, mass, adj_article, 
         article, adj_token_id, noun_token_id) %>%
  left_join(mean_ratings, by = c("adjective", "noun")) %>%
  rename(mean_rating = mean_typ) %>%
  filter(!is.na(mean_rating))


# NOTE: actual utterances from LDP are redacted for data sharing 
  # to preserve anonymity while sharing as much as possible
ldp_coded <- all_corpora %>%
  filter(dataset %in% c("ldp_parent_utts_udpipe", "ldp_child_utts_udpipe")) %>%
  left_join(orig_tokens, by = c("adj_token" = "adj", "noun_token" = "noun_orig")) %>%
  mutate(adjective = adj_token) %>%
  select(doc_id, noun_token, adjective, noun, prenominal, 
         dataset, mass, adj_article,adj_token, 
         article, adj_token_id, noun_token_id, subject, session, person) %>%
  left_join(mean_ratings, by = c("adjective", "noun")) %>%
  rename(mean_rating = mean_typ) %>%
  mutate(age = (4*session + 10)) %>%
  filter(!is.na(mean_rating))
```

```{r}
ldp_coded %>%
  filter(person == "parent") %>%
  mutate(corpus = "Parent speech") %>%
  select(adjective, noun, mean_rating, corpus) %>%
  bind_rows(cabnc %>% mutate(corpus = "Adult-adult speech") %>% 
              select(adjective, noun, mean_rating, corpus)) %>%
  mutate(typicality=mean_rating) %>%
  ggplot(aes(x = typicality)) +
  geom_density(fill="cornsilk") +
  xlab("More Atypical                   More Typical \n Typicality of adjective-noun pairs") +
  geom_vline(xintercept = 4, size=1, linetype="solid")+
  facet_wrap(~corpus, ncol = 1) +
  theme_few() +
  theme(#panel.grid = element_line(color="lightgrey",size=0.5), 
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(),
    axis.text.x = element_text(size=11, angle=28, hjust=1),
    axis.text.y = element_text(size=11),
    legend.position = "none") +
  scale_x_continuous(minor_breaks = seq(1 , 7, 1), breaks = seq(1, 7, 1), labels = c('never', 'rarely', 'sometimes', 'about half', 'often', 'almost always', 'always')) +
  scale_y_continuous(minor_breaks = seq(14, 58, 4), breaks = seq(14, 58, 4)) +
  theme(
    strip.background =element_rect(fill="transparent"),
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"), # get rid of legend panel bg
    legend.key = element_rect(fill = "transparent", colour = NA), # get rid of key legend fill, and of the surrounding
    axis.line = element_line(colour = "black") # adding a black line for x and y axis
)
```
```{r distribution-plot, fig.align = "center", fig.width=3.4, fig.height = 5.4, fig.cap = "Density plots showing parents' use of atypical and typical adjective-noun pairs across their child's age.", fig.pos="tb"}
ldp_coded %>%
  filter(person == "parent") %>%
  mutate(typicality=mean_rating) %>%
  group_by(session) %>%
  mutate(age = min(age)) %>%
  ggplot(aes(x = typicality, y=age, group=age, fill=age)) +
  geom_density_ridges2() +
  ylab("Child Age (months)") +
  xlab("More Atypical                   More Typical \n Typicality of adjective-noun pairs") +
  geom_vline(xintercept = 4, size=1, linetype="solid")+
  scale_fill_gradient(low="cornsilk", high=muted("red")) +
  theme_few() +
  theme(#panel.grid = element_line(color="lightgrey",size=0.5), 
    axis.line = element_line(colour = "black"),
    axis.ticks = element_line(),
    axis.text.x = element_text(size=11, angle=28, hjust=1),
    axis.text.y = element_text(size=11),
    legend.position = "none") +
  scale_x_continuous(minor_breaks = seq(1 , 7, 1), breaks = seq(1, 7, 1), labels = c('never', 'rarely', 'sometimes', 'about half', 'often', 'almost always', 'always')) +
  scale_y_continuous(minor_breaks = seq(14, 58, 4), breaks = seq(14, 58, 4)) +
  theme(
    strip.background =element_rect(fill="transparent"),
    panel.background = element_rect(fill = "transparent"), # bg of the panel
    plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
    panel.grid.major = element_blank(), # get rid of major grid
    panel.grid.minor = element_blank(), # get rid of minor grid
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"), # get rid of legend panel bg
    legend.key = element_rect(fill = "transparent", colour = NA), # get rid of key legend fill, and of the surrounding
    axis.line = element_line(colour = "black") # adding a black line for x and y axis
)
```



```{r compute_prototypicals}
parent_typical_ratings <- ldp_coded %>%
  filter(person == "parent") %>%
  group_by(age, adjective, noun, mean_rating) %>%
  count() %>%
  ungroup() %>%
  mutate(typical = mean_rating >= 5)

typicals_parent <- ldp_coded %>%
  filter(person == "parent") %>%
  group_by(age, adjective, noun, mean_rating) %>%
  count() %>%
  ungroup() %>%
  mutate(typical = mean_rating >= 5) %>%
  group_by(age, typical) %>%
  summarise(weighted_sum = sum(n), sum = n()) %>%
  pivot_longer(cols = c(sum, weighted_sum), names_to = "measure", values_to = "sum") %>%
  group_by(age, measure) %>%
  mutate(prop = sum / sum(sum)) %>%
  filter(typical)

typicals_cabnc <- cabnc %>%
  group_by(adjective, noun, mean_rating) %>%
  count() %>%
  ungroup() %>%
  mutate(typical = mean_rating >= 5) %>%
  group_by(typical) %>%
  summarise(weighted_sum = sum(n), sum = n()) %>%
  pivot_longer(cols = c(sum, weighted_sum), names_to = "measure", values_to = "sum") %>%
  group_by(measure) %>%
  mutate(prop = sum / sum(sum)) %>%
  filter(typical)


typical_type_lmer <- glmer(typical ~ log(age) + (1|noun), 
                      data = parent_typical_ratings, 
     family = "binomial") %>% 
  tidy() %>%
  filter(effect == "fixed")

typical_token_lmer <- glmer(typical ~ log(age) + (1|noun), 
                      data = parent_typical_ratings, 
                      weights = n,
     family = "binomial") %>% 
  tidy() %>%
  filter(effect == "fixed")
```
```{r typical-effects}
typical_effect <- typical_token_lmer %>% filter(term == "log(age)") %>%
  pull(estimate)
typical_statistic <- typical_token_lmer %>% filter(term == "log(age)") %>% 
  pull(statistic)
typical_p <- typical_token_lmer %>% filter(term == "log(age)") %>% 
  pull(p.value) %>% printp()
```

```{r generics}
generics_cb <- read_csv(here("../data/generics_only_cb.csv")) %>%
  rename(generics_manual_cb = generics_manual,
         generics_adj_noun_cb = generics_adj_noun,
         generics_neg_question_cb = generic_neg_question)

generics_bm <- read_csv(here("../data/generics_only_bm.csv")) %>%
  rename(generics_manual_bm = generics_manual,
         generics_adj_noun_bm = generics_adj_noun,
         generics_neg_question_bm = negation_or_question)

generics_all <- generics_cb %>% left_join(generics_bm)

generics_agreement <- generics_all %>%
  filter(!is.na(generics_manual_cb) & !is.na(generics_manual_bm)) %>%
  filter(generics_manual_cb == generics_manual_bm) %>% nrow()/110

generics_agreed <- generics_all %>%
  filter(!is.na(generics_manual_cb) & !is.na(generics_manual_bm)) %>%
  filter(generics_manual_cb == 1 & generics_manual_bm == 1) %>% nrow()

generics_agreement_adj_noun <- generics_all %>%
  filter(!is.na(generics_manual_cb) & !is.na(generics_manual_bm)) %>%
  filter(generics_manual_cb == 1 & generics_manual_bm == 1) %>%
  filter(generics_adj_noun_bm == generics_adj_noun_cb) %>% nrow()/generics_agreed

generics_disagreed <- generics_all %>%
  filter(!is.na(generics_manual_cb) & !is.na(generics_manual_bm)) %>%
  filter(generics_manual_cb != generics_manual_bm) 

generics_consensus <- generics_all %>%
  mutate(generics_consensus = generics_manual_cb | generics_manual_bm,
         generics_adj_noun_consensus = generics_adj_noun_bm | generics_adj_noun_cb,
         generics_neg_question_consesus = generics_neg_question_bm | generics_neg_question_cb) %>%
  mutate(generics_consensus = if_else(is.na(generics_consensus), FALSE, generics_consensus),
         generics_adj_noun_consensus = if_else(is.na(generics_adj_noun_consensus), FALSE, generics_adj_noun_consensus),
         generics_neg_question_consesus = if_else(is.na(generics_neg_question_consesus), FALSE, generics_neg_question_consesus))

total_utterances <- generics_consensus %>%
  distinct(doc_id, sentence)

total_generics <- generics_consensus %>%
  filter(generics_consensus) %>%
  distinct(doc_id, sentence)

total_generics_adj_noun <- generics_consensus %>%
  filter(generics_consensus & generics_adj_noun_consensus) %>%
  distinct(doc_id, sentence)

total_generics_no_neg_q <- generics_consensus %>%
  filter(generics_consensus & generics_adj_noun_consensus & !generics_neg_question_consesus) %>%
  distinct(doc_id, sentence)

total_generics_neg_q <- generics_consensus %>%
  filter(generics_consensus & generics_adj_noun_consensus & generics_neg_question_consesus) %>%
  distinct(doc_id, sentence)


generics_analysis <- generics_consensus %>%
  filter(generics_consensus & generics_adj_noun_consensus & !generics_neg_question_consesus) %>%
  rename(adjective = adj_lower, noun = noun_stem) %>%
  left_join(mean_ratings) %>%
  rename(mean_typ_1 = mean_typ) %>%
  rename(noun_stem = noun, noun = noun_lemma) %>%
  left_join(mean_ratings) %>%
  rename(mean_typ_2 = mean_typ) %>%
  mutate(mean_typ = if_else(is.na(mean_typ_1), mean_typ_2, mean_typ_1)) %>%
  select(-mean_typ_1, -mean_typ_2)

mean_generic_rating = mean(generics_analysis$mean_typ)

hist(generics_analysis$mean_typ)

t.test()
```

On a subset of 110 utterances coded by both raters, raters had `r generics_agreement`% agreement about whether an utterance was a generic. Among that subset, they had `r generics_agreement_adj_noun`% agreement about whether the coded adjective-noun pair was described generically. For utterances coded by both raters, in all subsequent statistics, we conservatively took the more lenient code (i.e., coding an utterance as a generic) from the two raters.

Among the 448 utterances identified by the model as potentially being generics, 180 were manually verified by human raters as being true generics. Among these, the majority of generics did not describe a generic property that predicated the target adjective-noun pair. For example, many generics had the adjective-noun pair in an embedded noun phrase (e.g., "People love orange juice", "The red light means stop"). Only 28 utterances were coded as both being generic and predicating the relevant adjective-noun pair (e.g., "Floors are dirty", "Strawberries are red"). Five of these were flagged by raters as being negated or questioning the generic statement (e.g., "Roses are red?", "Ducks are not furry") and excluded on this basis. This leaves a set of `r nrow(generics_analysis)` generics that predicate the relevant adjective-noun pair.

Examining the typicality of adjective-noun pairs predicated in generics, the features expressed in generics were on average more typical than those in the full set of utterances (M = `r mean_generic rating`).


